---
title: "phytotraitr"
output:
  html_document:
    keep_md: yes
  pdf_document: default
---
*********************
This package contains a compilation of compilations of published phytoplankton trait data. 

## Contents:

* Nutrient uptake traits from Edwards et al (2015)
* Thermal traits from Chen (2015)
* Cell and natural unit biovolumes from Kremer et al (2014)


[Chen B. (2015) Patterns of thermal limits of phytoplankton. Journal of Plankton Research 37, 285-292.](http://www.plankt.oxfordjournals.org/cgi/doi/10.1093/plankt/fbv009)

[Edwards K.F., Klausmeier C.A. & Litchman E. (2015) Nutrient utilization traits of phytoplankton. Ecology 96, 2311-2311.](http://www.esajournals.org/doi/abs/10.1890/14-2252.1)

[Kremer C.T., Gillette J.P., Rudstam L.G., Brettum P. & Ptacnik R. (2014) A compendium of cell and natural unit biovolumes for >1200 freshwater phytoplankton species. Ecology 95, 2984.](http://www.esajournals.org/doi/abs/10.1890/14-0603.1)

## Installation

phytotraitr is not available from CRAN.

You can install directly from github with:
```R
# install.packages("devtools")
devtools::install_github("andrewdolman/phytotraitr")
```

## Usage

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(phytotraitr)
```

Each table has a corresponding *_key table with details about each variable. The first three columns of the keys are also found in the help files for each table.

```{r tables, results='markup'}
kable(Edwards_nutrient_traits[1:10,1:6], format = "markdown")
kable(Edwards_nutrient_traits_key[1:6,], format = "markdown")
```

### Minimum and maximum P cell quotas for freshwater taxa

```{r min_max_P_C}
p <- Edwards_nutrient_traits %>% 
  filter(system == "freshwater") %>%
  select(taxon, qmin_p_c, qmax_p_c) %>% 
  gather(Trait, Value, -taxon) %>% 
  filter(complete.cases(Value)) %>% 
  ggplot(aes(x = taxon, y = Value, colour = Trait)) %>% 
  + geom_point(position = position_jitter(width = 0.05)) %>% 
  + coord_flip() %>% 
  + scale_x_discrete(with(Edwards_nutrient_traits_key,
                          Edwards_nutrient_traits_key[`Column name`=="taxon","Variable definition"])) %>% 
  + scale_y_continuous(paste0("[",
                              with(Edwards_nutrient_traits_key,
                                   Edwards_nutrient_traits_key[`Column name`=="qmin_p_c","Units"],
                              "]")))
p
```


## Consistency

The data from the separate publications have not been harmonized - i.e. variable names may differ (e.g. species, Species) and they may use different taxonomy.


## Full list of dataframes

```{r}
sets <- data(package = "phytotraitr")  
sets <- data.frame(sets[[3]])  %>% 
  select(Item, Title)

kable(sets, format = "markdown")
```

